Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-12-21 supplement.tex
added text about initialization

For the early history, we tested model parameterizations that cover many of the
proposed scenarios of population structure, size changes, and/or archaic
hominin admixture. The simplest model, in terms of number of parameters, was a
single-origin expansion of humans, with no structure in the stem and no archaic
hominin admixture aside from the Neanderthal admixture in Eurasian populations
following the out-of-Africa migration. This model allowed for a population size
change in the stem of humans between the ancestral split of the
human-Neanderthal lineages and the more recent split of branches leading to
southern and western/eastern African populations. The recent expansion and the
African multi-regional models (Figure~\ref{fig:proposed-models}A and D) have
the same topology, so interpretation of the model depends on the specified or
inferred divergence times. Thus, because model optimization explores a wide
range of parameter values, we account for both scenarios with this
parameterization.

>>>>

As is commonly done in demographic inference, 
all models we considered are initialized from an ancestral panmictic population at 
mutation-recombination-drift equilibrium. This allows for rapid initialization of the 
summary statistics to reasonable values.
The size $N_e$ of this ancestral population is a model-specific adjustable parameter 
that also serves as a reference size to define subsequent population size changes.

For the early history following this initialization, 
we tested model parameterizations that cover many of the
proposed scenarios of population structure, size changes, and/or archaic
hominin admixture. The simplest model, in terms of number of parameters, was a
single-origin expansion of humans, with no structure in the stem and no archaic
hominin admixture aside from the Neanderthal admixture in Eurasian populations
following the out-of-Africa migration. This model allowed for a population size
change in the stem of humans between the ancestral split of the
human-Neanderthal lineages and the more recent split of branches leading to
southern and western/eastern African populations. The recent expansion and the
African multi-regional models (Figure~\ref{fig:proposed-models}A and D) have
the same topology, so interpretation of the model depends on the specified or
inferred divergence times. Thus, because model optimization explores a wide
range of parameter values, we account for both scenarios with this
parameterization.
